Multi-stage programming with functors and monads: eliminating abstraction overhead from generic code

### 2007-07-09 SCP2008/paper/metamonads.tex
finish the first sub-section describing the aspects

@@

@@

\noindent \jacques{Absorb this in the above} These are separated out from the
others as they are cross-cutting concerns: in the case of the length measure, a
property of the domain will influence the pivoting method \emph{if} pivoting is
to be performed.\jacques{Also, should mention a few more aspects like
inplace, errors-on-singular, logging, sparsity, as things that are not
implemented}

>>>>

\newcounter{naspects}

@@

\setcounter{naspects}{\value{enumi}}

@@

There are other aspects which we have not (yet) implemented, but that have
been seen in different implementations.  In particular,
\begin{enumerate}
\setcounter{enumi}{\value{naspects}}
\item \textbf{Logging}: This is a classical cross-cutting concern, which
is also useful in this context.  Of course, there are many different
parts of the algorithm which could be ``logged'', and ensuring fine control
over this is a thorny issue.
\item \textbf{Sparsity}: If a matrix is known to be sparse, the traversal
(at least) should be sparse.  Additionally, it would better still if
care was taken to minimize fill-in.
\item \textbf{Other structure}: For example, if a matrix is known in
advance to be real symmetric tri-diagonal, then LU decomposition can be
done in $O(n^2)$ time (instead of $O(n^3)$; the main difficulty in 
implementing this kind of reduction is the fact that this version of 
LU requires $O(n)$ extra storage (unlike all other versions).
\item \textbf{Warnings}: In the case of LU decomposition of a matrix over
a domain for which we only have a heuristic algorithm for zero
testing, it is customary to issue a warning (or otherwise log) when 
a potentially zero pivot is chosen.
\item \textbf{In-place}: When it is possible, it would be good to have the
option to re-use the input matrix as the storage space for the output.
\item \textbf{Error-on-singular}: There are instances where we are only
interested in an LU decomposition for a non-singular matrix, and would
prefer to get an exception when a singular matrix is encountered.
\setcounter{naspects}{\value{enumi}}
\end {enumerate}

Clearly the above $\thenaspects$ aspects have a rather complex 
dependency graph.  While some aspects are purely orthogonal to others, 
most are inter-dependent.  For example, if one wants the determinant of
the matrix as part of the output, then the determinant has to be tracked
during the algorithm; a priori, it is possible to make incompatible choices,
and our implementation has to insure that this is in fact impossible.
More precisely, our goal is to make sure that this is detected long
before the algorithm could be run.

There are other kinds of dependencies: for example, whether or not a 
domain has a length measure will influence pivoting, \emph{if} pivoting
is to be performed.  While there are no circular dependencies, there is
no clear directed graph of dependencies.  In other words, one cannot 
simply use product, co-products, and compositions of the various aspects 
to build the resulting algorithm.  We will get back to this point when
we describe the details of the implementation.
